---
output:
  word_document: default
  html_document: default
---
##Regresion Project

#Setting up the environment
```{r}
rm(list=ls())
library(readxl)
library(rio)
library(car)
```

#Importing Dataset

```{r}
population=import("/Users/siddharthrawat/Downloads/6304\ Regression\ Project\ Data.csv")
colnames(population)=tolower(make.names(colnames(population)))
attach(population)
#converting payment_type to factor variable
population$payment_type = as.factor(population$payment_type)
```

#Setting seed & creating sample

```{r}
set.seed(45775355)
sample.taxi.data=population[sample(1:nrow(population),100,replace=FALSE),]
names(sample.taxi.data)
nrow(sample.taxi.data)
ncol(sample.taxi.data)
```

#cleaning the sample data

```{r}
clean.taxi.data = subset(sample.taxi.data, trip_seconds>0)
nrow(clean.taxi.data)
ncol(clean.taxi.data)
subset.taxi.data = subset(clean.taxi.data, select = c(trip_seconds,trip_miles,fare,tips, extras,tolls,trip_total))
summary(subset.taxi.data)
```

#Summary & Density Plots

```{r}
attach(subset.taxi.data)
plot (density(trip_seconds))
plot (density(trip_miles))
plot (density(fare))
plot (density(tips))
plot (density(tolls))
plot (density(extras))
plot (density(trip_total))
#dropping tolls
subset.taxi.data = subset(clean.taxi.data, select = c(trip_seconds,trip_miles,fare,tips, extras,trip_total))
```


#Correlation analysis of the continuous variables

```{r}
summary(clean.taxi.data$payment_type)
```

#Correlation analysis of the continuous variables

```{r}
#Select a subset of numeric variables for correlation analysis
plot(subset.taxi.data)
cor(subset.taxi.data)
```

#Correlation Matrix

```{r}
library(corrplot)
xx=cor(subset.taxi.data)
corrplot(xx,method="number")
```


#Initial Regression Model

```{r}
taxi.out=lm(fare~trip_seconds+trip_miles+payment_type, data = clean.taxi.data)
summary(taxi.out)
confint(taxi.out)
```

#Kitchen Sink Model

```{r}
clean.taxi.data$trip_seconds2=clean.taxi.data$trip_seconds^2
clean.taxi.data$trip_miles2=clean.taxi.data$trip_miles^2
clean.taxi.data$trip_seconds3=clean.taxi.data$trip_seconds^3
clean.taxi.data$trip_miles3=clean.taxi.data$trip_miles^3
clean.taxi.data$inter=clean.taxi.data$trip_seconds * clean.taxi.data$trip_miles
taxi.out2=lm(fare~trip_seconds+trip_miles+trip_seconds2+trip_miles2+trip_seconds3+trip_miles3+payment_type+inter+tips+extras, data = clean.taxi.data)
summary(taxi.out2)
```

#Stepwise Regression

```{r}
step(lm(fare~trip_seconds+trip_miles+trip_seconds2+trip_miles2+trip_seconds3+trip_miles3+payment_type, data = clean.taxi.data, direction = "both"))
```

#finalreg model

```{r}
taxi.finalreg=lm(fare~trip_seconds+trip_seconds2+trip_miles2, data = clean.taxi.data)
summary(taxi.finalreg)
```

#LINE Conformity

```{r}
#Linearity
new_subset=subset(clean.taxi.data,select=c("trip_seconds","trip_seconds2","trip_miles2"))
plot(new_subset)

#Normality
qqnorm(taxi.finalreg$residuals,pch=19)
qqline(taxi.finalreg$residuals,lwd=3,col="red")

#Equality of Variances
plot(taxi.finalreg$fitted.values,rstandard(taxi.finalreg),pch=19)
abline(0,0,col="red",lwd=3)
```

#Leverage Points

```{r}
lev=hat(model.matrix(taxi.finalreg))
plot(lev,pch=19)
abline(3*mean(lev),0,col="red",lwd=3)
```

#Finding the Outlier

```{r}
outliers = clean.taxi.data[lev>(3*mean(lev)),]
outliers
```

#Removing the outliers

```{r}
reduced.taxi.data=clean.taxi.data[-which(clean.taxi.data$taxi_id %in% outliers$taxi_id & clean.taxi.data$trip_seconds!=300),]
```

#And rerun the regression

```{r}
taxi.redreg=lm(fare~trip_seconds+trip_seconds2+trip_miles2, data = reduced.taxi.data)
summary(taxi.redreg)
stdresids=rstandard(taxi.redreg)
#Linearity
plot(reduced.taxi.data$fare,taxi.redreg$fitted.values,
     pch=21,main="Actual Fare v. Fitted")
abline(0,1,lwd=3,col="red")

#Normality
qqnorm(taxi.redreg$residuals,pch=21)
qqline(taxi.redreg$residuals,lwd=3,col="red")

#Equality of Variances
plot(taxi.redreg$fitted.values,rstandard(taxi.redreg),pch=21)
abline(0,0,col="red",lwd=3)

```

#Rerun the Model with new sample

```{r}
set.seed(45775360)
sample.taxi.data2=population[sample(1:nrow(population),100,replace=FALSE),]
names(sample.taxi.data2)
```

#cleaning the data
```{r}
clean.taxi.data2 = subset(sample.taxi.data2, trip_seconds>0)
nrow(clean.taxi.data2)
ncol(clean.taxi.data2)
subset.taxi.data2 = subset(clean.taxi.data2, select = c(trip_seconds,trip_miles,fare,tips, extras,trip_total))
```

#finalreg model 2
```{r}
clean.taxi.data2$trip_seconds2=clean.taxi.data2$trip_seconds^2
clean.taxi.data2$trip_miles2=clean.taxi.data2$trip_miles^2
taxi.finalreg2=lm(fare~trip_seconds+trip_seconds2+trip_miles2, data = clean.taxi.data2)
summary(taxi.finalreg2)
```
```{r}
vif(taxi.finalreg2)
vif(taxi.finalreg)
```